# Results and Discussions

```{r setup2, file = "R/chapter_start.R", include = FALSE, cache = FALSE}
library(plm)
library(sandwich)
```

```{r model-setups, include=FALSE}
f <- function(x) format(round(x, 3), big.mark=",")
gm <- list(
  list("raw" = "nobs", "clean" = "No. of Obs.", "fmt" = f),
  list("raw" = "logLik", "clean" = "Log. Liklihood", "fmt" = f),
  list("raw" = "aic", "clean" = "AIC", "fmt" = f),
  list("raw" = "r.squared", "clean" = "R\u00b2", "fmt" = f))

cm <- c(
  "sev_day_avg" = "Sev-Day No. of Acts.",
  "sexFemale" = "Female", 
  "age" = "Age", 
  "fsiq_2" = "IQ Score", 
  "prescribed_groupAutism" = "Autism", 
  "prescribed_groupSocial Anxiety" = "Social Anxiety", 
  "suicidal_ideation_q31_evenTRUE" = "Suicidal Ideation",
  "energy" = "Energy",
  "motivation" = "Motivation", 
  "area" = "Activity Area",
  "length" = "Distance", 
  "numTrips" = "No. of Activities")

cm_comp <- c(
  "numTrips" = "No. of Activities",
  "log(numTrips + 0.1)" = "log (No. of Activities)",
  "I(numTrips^2)" = "No. of Activities\u00b2",
  "sev_day_avg" = "Sev-Day No. of Acts.",
  "area" = "Activity Area",
  "length" = "Distance",
  "log(length)" = "log (Distance)",
  "I(length^2)" = "Distance\u00b2",
  "suic_intensity" = "Suicidal Intesity",
  "park" = "Park",
  "library" = "Library",
  "grocery" = "Grocery Store",
  "social_rec" = "Social Recreation", 
  "sev_day_avg_park" = "Seven-Day Park",
  "sev_day_avg_grocery" = "Seven-Day Grocery",
  "sev_day_avg_library" = "Seven-Day Library")

cm_group <- c(
  "numTrips" = "No. of Activities",
  "sev_day_avg" = "Sev-Day No. of Acts.",
  "park" = "Visits to Parks",
  "grocery" = "Visits to Grocery Store",
  "library" = "Visits to Libraries",
  "social_rec" = "Visits to Social Recreation Locations")
```

Participants in this study belong to one of three groups: individuals with a psychiatric diagnosis of social anxiety disorder, individuals diagnosed with autism spectrum disorder (ASD), or individuals in a designated control group.
Analyzing travel behavior and mental well-being in this study requires a nuanced understanding of each participant group, including how their travel patterns differ and how these patterns influence their overall well-being.
We aim to understand differences in the number of activities each group engages in, differences in motivation levels across groups, and the types of activities that impact motivation for the three group, in order to better understand the relationship between travel behavior and mental well-being.

## Activity Engagement by Group

Examining the number of activities engaged in by each group is important because it offers insight into overall mobility, levels of social participation, and engagement with the environment, which are all closely tied to mental well-being.
Reduced activity levels may reflect barriers to access, lower motivation, or avoidance behaviors commonly associated with mental health conditions such as ASD and social anxiety.
By understanding how activity frequency varies between groups, we can begin to identify patterns that highlight how different populations experience and navigate the world around them, which in turn helps clarify how travel behavior influences mental health outcomes.

We analyzed the total number of activities and the seven-day rolling average number of activities for individuals comprising the three different groups.
The descriptive statistics in @tbl-groupActs describe these results for each group.

```{r table groupAct, echo=FALSE, message = FALSE, warning=FALSE}
#| label: tbl-groupActs
#| tbl-cap: "Activity Descriptive Statistics by Group"

tar_load(model_data)

acts <- model_data %>% 
  filter(!is.na(numTrips)) %>%
  select(prescribed_group, numTrips, sev_day_avg) %>% 
# , park, grocery, library, social_rec) %>% 
  group_by(prescribed_group) %>% 
  rename("No. of Activities" = numTrips,
         "Sev-Day No. of Acts." = sev_day_avg)

         # "Visits to Parks" = park,
         # "Visits to Grocery Store" = grocery,
         # "Visits to Libraries" = library,
         # "Visits to Social Recreation Locations" = social_rec)

datasummary_balance(~prescribed_group, data = acts, dinm=FALSE, align = "lcccccc", output = "kableExtra") |> 
  kable_styling(latex_options="scale_down")
```

The data reveal clear differences in activity level across groups.
Individuals in the control group engaged in an average of 3.0 activities, compared to 2.6 for the social anxiety group and 2.1 for the autism group.
The seven-day rolling average number of activities showed similar trends but with slightly lower variability.
These patterns suggest that autistic individuals or those with social anxiety may face greater barriers to engaging in activities or have lower motivation compared to those in the control group.
This underscores the importance of considering group differences when analyzing activity patterns and mental health outcomes and by examining these group differences, we can better understand how travel behavior intersects with mental well-being.

## Motivation by Group

Motivation, as a proxy for overall well-being, reflects individuals' energy, engagement, and readiness to participate in daily trips and activities as motivation plays a role in both initiating and sustaining participation in activities.
Based on existing literature, we expected that individuals in the autism and social anxiety groups would report lower levels of motivation, or well-being, compared to the control group.
Investigating differences in motivation levels across groups is essential for understanding how internal psychological factors influence travel behavior and activity engagement.
Lower motivation levels may indicate underlying challenges that may be more common among autistic individuals and those with social anxiety.
By comparing motivation levels across groups, we can gain deeper insights into how mental health conditions affect an individual’s drive to engage with their surroundings, ultimately shedding light on the complex relationship between mental well-being and mobility patterns.

Using motivation as an indicator of well-being, we analyzed motivation levels from the evening survey, rated on a 1-100 scale.
We observed notable differences in motivation levels across the groups.
The control group had the highest mean motivation score of 47.7, falling slightly below the middle range of the "typical motivation" category.
The autism group had the lowest mean score of 34.0, falling in the "enough motivation to get by" category.
The social anxiety group had a mean score of 40.3, falling between the other two groups and on the lowest end of the "typical motivation" category.

These results highlight significant differences in motivation levels, with the control group showing higher motivation than both the autism and social anxiety groups, and the social anxiety group showing higher motivation than the autism group.
These findings align with expectations based on existing literature and highlight the nuanced ways in which mental health conditions can affect day-to-day well-being.
Understanding these disparities is crucial for interpreting how individuals engage with their environments and how travel behavior may either support or hinder well-being across different populations.

## Building and Evaluating Models

To explore the effect of number of activities individuals participated in on motivation levels, we ran ordinary least squares (OLS), fixed effects (FE), and random effects (RE) models, originally not taking into account the group associations, to determine the most appropriate model approach going forward.
Robust standard errors and t-statistics were used to account for potential autocorrelation and heteroskedasticity in the panel data.
The results of these three models are shown in @tbl-olsfere.

```{r table ols fe re, message=FALSE, echo=FALSE, warning=FALSE}
#| label: tbl-olsfere
#| tbl-cap: "OLS, FE, and RE Models (Y-value: motivation level)"
#| warnings: FALSE
 
library(plm) 

tar_load(ols)
tar_load(fe)
tar_load(re)

models_olsfere <- list(
    "OLS" = ols,
    "FE" = fe,
    "RE" = re
   )

modelsummary(models_olsfere, gof_map = gm, coef_map = cm,
              vcov = function(x) vcovHC(x, cluster = "time"),
              estimate = c("{estimate}{stars}"),
              statistic = c("({statistic})"),
              notes = list("Robust t-statistics in parentheses. + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001"),
              output = "kableExtra"
   )
```

To determine the most appropriate modeling approach, we used the Hausman test to compare the RE and FE models.
The Hausman test evaluates whether RE estimates are consistent and efficient relative to FE estimates.
The null hypothesis assumes that the RE model is appropriate, while the alternative favors the FE model.
With a p-value of 0.0013, well below the 0.05 threshold, we reject the null hypothesis, indicating that the RE model is inconsistent.
As a result, we selected the FE model for the remainder of the analysis.

### Effect of Demographic Factors on Motivation

Given the results of the Hausman test, we proceeded with the FE model to analyze the activity pattern and mental health data.
While the FE model effectively controls for individual-specific unobserved heterogeneity, a key limitation is its inability to estimate the effects of time-constant or time-invariant variables.
To address this gap, we conducted a linear regression using the intercepts from the FE model as the dependent variable.
This linear regression is described generally in @eq-linreg

$$
\bar{y}_{i} \; \tilde{} \; \beta (\vec{X}_{it})
$$ {#eq-linreg}

This allowed us to examine how demographic characteristics, such as sex, age, IQ score, and group association, are associated with baseline motivation levels across individuals.
The results from this model are shown in @tbl-fedemolm.

```{r table fe demo lm, message=FALSE, echo=FALSE, warning=FALSE}
#| label: tbl-fedemolm
#| tbl-cap: "Fixed Effects and Demographics Regression (Y-value: motivation level)"

tar_load(fe_model)

modelsummary(fe_model, gof_map = gm, coef_map = cm,
             estimate = c("{estimate} ({statistic}){stars}"),
             statistic = NULL,
             notes = list("t-statistics in parentheses. + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001"),
             output = "kableExtra"
             )
```

The linear regression analysis revealed several significant associations related to motivation levels.
Female participants, on average, had motivation scores 6.351 points lower than their male counterparts.
Age and IQ score were not significantly associated with motivation.
However, group association showed strong effects: autistic individuals scored 10.247 points lower in motivation compared to the control group, while those with social anxiety showed an 8.544-point decrease.
These findings suggest that sex and group status are important factors influencing baseline motivation levels, whereas age and IQ play a more limited role.
The model accounted for approximately 26.4% of the variance in motivation, indicating that additional variables may also contribute to motivation differences.
Overall, these results underscore the importance of accounting for individual differences, particularly sex and group association, when examining motivation, with a focus on group typology for further exploration.

### Need to Account for Group and Individual Differences

To accurately examine the relationship between travel behavior and mental well-being, it is important to account for differences both between and within groups.
Given the statistically significant differences in average number of activities and motivation levels across the three groups, it was essential to model each group separately.
This approach allows us to better account for the unique behavioral patterns, needs, and challenges associated with each group, particularly those stemming from differences in mental health conditions.
Additionally, using FE models enables us to control for unobserved individual-level factors that may influence motivation, ensuring that variations within each person over time are properly considered.
By analyzing each group independently and leveraging individual-level FE, we can more accurately explore the relationship between travel behavior and well-being while recognizing the distinct experiences and baselines of participants within each subgroup.

To visualize the need account for difference both between and within groups, we plotted the number of activities and motivation levels, on a given activity-day, for all individuals within each group.
@fig-groupFE presents the results of the FE models for all three groups illustrating the relationship between motivation and the number of activities by group.
The "within" lines represent FE model estimates that account for individual differences, modeling how changes in activity levels relate to motivation *within each person* over time given their baseline motivation levels.
The "pooling" lines, by contrast, depict OLS estimates, modeling the relationship without accounting for individual variation.
The divergence between the within and pooling lines highlights the importance of accounting for individual differences within each group when examining behavior and mental health outcomes.
Notably, the autism and social anxiety groups show steeper slopes when all data are pooled together; however, when individuals are analyzed separately, the slopes for each group are much less steep.
These results align more closely with expectations from existing literature.

```{r figure groupFE, message=FALSE, echo=FALSE, warning=FALSE}
#| label: fig-groupFE
#| fig-cap: "FE model for motivation vs. number of activities by group."

## ALL THREE GROUPS
 fixed <- plm(motivation ~ numTrips, index = c("userId", "activityDay"), data = model_data)
 
z <- fixef(fixed) 
fe <- tibble(fe = z, userId = names(z))
data_f <- model_data |>
  as_tibble() |>
  transmute(
    userId, activityDay, prescribed_group, motivation, numTrips
  ) %>% 
  left_join(fe, by = "userId") %>% 
  filter(!is.na(fe) & !is.na(numTrips) & !is.na(motivation))

userId_mean <- data_f %>% 
  group_by(userId, prescribed_group) %>% 
  summarise(
    motivation = mean(motivation),
    numTrips = mean(numTrips),
    fe = mean(fe)
  )

segw = 2
ggplot() + 
  geom_point(data = data_f, aes(x = numTrips, y = motivation), alpha = 0.2, color = "gray30") + 
  geom_smooth(data = data_f, aes(x = numTrips, y = motivation, linetype = "Pooling"), method = "lm", se = FALSE, color = "black", linewidth = 1) + 
  geom_point(data = userId_mean, aes(x = numTrips, y = motivation, color = prescribed_group, fill = prescribed_group), shape = 21, size = 1.5) +
  geom_segment(data = userId_mean, aes(x = numTrips - segw, xend = numTrips + segw,
                                       y = motivation - segw * fixed$coefficients[1],
                                       yend = motivation + segw * fixed$coefficients[1],
                                       color = prescribed_group, linetype = "Within"), linewidth = 0.6) +
  facet_wrap(~prescribed_group) +
  guides(color = FALSE, fill = FALSE) +
  theme_bw() +
  xlab("Number of Activities") +
  ylab("Level of Motivation") +
  scale_linetype_manual(name = "Legend", values = c("Pooling" = "dashed", "Within" = "solid")) +
  theme(legend.position = "right") +
  scale_x_continuous(breaks = seq(min(data_f$numTrips, na.rm = TRUE), max(data_f$numTrips, na.rm = TRUE), by = 2)) +
  scale_y_continuous(breaks = seq(min(data_f$motivation, na.rm = TRUE), max(data_f$motivation, na.rm = TRUE), by = 20))
```

Since the FE models require both an evening survey response for motivation and the number of activities determined by the DBSCAN-TE algorithm, some data were lost.
This reduced the sample size from 31 to 23 in the control group, from 29 to 17 in the autism group, and from 28 to 22 in the social anxiety group.
The FE models need enough data points for each participant to ensure effective analysis, which explains the reduction in participants.

### Fixed Effects Models for Motivation and Number of Activities

After visualizing the relationships, we proceeded and constructed FE models separately for each group to better understand how motivation levels relate to the number of activities engaged in for distinct groups of individuals.
The results from these models are presented in @tbl-groupFEAnaly.

```{r table groupAnaly, message=FALSE, echo=FALSE, warning=FALSE}
#| label: tbl-groupFEAnaly
#| tbl-cap: "FE Models: Motivation and Number of Activities by Group"

sep_groups <- model_data %>% 
  group_by(prescribed_group) %>% 
  nest() %>% 
  mutate(
    model = map(data, function(d) plm(motivation ~ numTrips, index = c("userId", "activityDay"), data = d, model = "within"))
  ) 

modelsummary(sep_groups$model %>% set_names(str_c("Group: ", sep_groups$prescribed_group)),
             gof_map = gm, coef_map = cm_comp,
             vcov = function(x) vcovHC(x, cluster = "time"),
             estimate = c("{estimate}{stars}"),
             statistic = c("({statistic})"),
             notes = list("Robust t-statistics in parentheses. + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001"),
             output = "kableExtra") 
```

The results revealed varying relationships across the control, autism, and social anxiety groups.
In the control group, the coefficient was 0.259, suggesting a weakly positive but not strongly significant association between the number of activities and motivation.
For the autism group, the coefficient was 0.361, also positive but not significant, indicating that increased activity may have limited direct impact on motivation in this group.
In contrast, the social anxiety group exhibited a coefficient of 0.483, which was statistically significant.
This suggests a meaningful positive relationship: as activity levels increase, motivation also increases among individuals with social anxiety.

These findings emphasize the value of modeling by group and accounting for individual differences.
While engaging in more activities appears to significantly benefit motivation levels for those with social anxiety, similar effects were not observed in the control or autism groups.
This may indicate that other factors might play a more critical role in influencing motivation levels for the control and autism group.

## Activity Types

To deepen our understanding of how activity engagement influences motivation, we next explored whether the *type* of location visited played a role.
Specifically, we examined how the number of activities at certain location categories—such as parks, grocery stores, and libraries—impacted motivation levels.
The analysis incorporated activity counts derived from the DBSCAN-TE algorithm, as well as seven-day and 14-day moving averages to capture short- and medium-term behavioral trends.

While all location types and time windows were considered, only a few yielded statistically significant results.
These findings highlight the importance of considering both the quantity and context of activities when analyzing travel behavior and mental well-being.

### Activities at Parks and Grocery Stores

Building on the general relationship between activity levels and motivation, we further investigated how specific types of activities at different locations influence motivation within each group.

When examining park-related activities, statistically significant results were found exclusively for the control group.
@tbl-groupFEParks presents the FE models for the seven-day rolling average number of activities at parks for the three groups.
Specifically, each additional park activity within a seven-day period was associated with a 3.017-point increase in motivation level.
This suggests that more frequent park visits are linked to heightened motivation among individuals without a psychiatric diagnosis.
However, no significant relationship was observed between park visits and motivation for individuals in the autism or social anxiety groups, indicating that park activities during the observed time frame may not meaningfully influence motivation for these populations.

```{r table groupParks, message=FALSE, echo=FALSE, warning=FALSE}
#| label: tbl-groupFEParks
#| tbl-cap: "FE Models: Motivation and Number of Activities at Parks by Group"

locations <- model_data %>% 
  select(userId, activityDay, prescribed_group, motivation, numTrips, park, sev_day_avg_park, library, sev_day_avg_library, grocery, sev_day_avg_grocery, social_rec) %>% 
  group_by(prescribed_group) %>% 
  nest() %>% 
  mutate(
    model = map(data, function(d) plm(motivation ~ sev_day_avg_park, index = c("userId", "activityDay"), data = d, model = "within"))
  ) 

modelsummary(locations$model %>% set_names(str_c("Group: ", locations$prescribed_group)),
             gof_map = gm, coef_map = cm_comp,
             vcov = function(x) vcovHC(x, cluster = "time"),
             estimate = c("{estimate}{stars}"),
             statistic = c("({statistic})"),
             notes = list("Robust t-statistics in parentheses. + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001"),
             output = "kableExtra")
```

In contrast, grocery store visits revealed more complex patterns across the groups.
@tbl-groupFEGrocery presents the FE models for the number of activities at grocery stores for the three groups.
For the autism group, a statistically significant negative correlation emerged: each additional grocery store activity was associated with a 2.725-point decrease in motivation.
A similar, though less significant, effect was found for the control group, with each additional daily grocery store activity linked to a 2.690-point drop in motivation.
The social anxiety group, however, did not show any statistically significant associations.
These results suggest that grocery store visits, particularly for autistic individuals, may be associated with increased stress or lower well-being, thus reducing motivation.

```{r table groupGrocery, message=FALSE, echo=FALSE, warning=FALSE}
#| label: tbl-groupFEGrocery
#| tbl-cap: "FE Models: Motivation and Number of Activities at Grocery Stores by Group"

locations <- model_data %>% 
  select(userId, activityDay, prescribed_group, motivation, numTrips, park, sev_day_avg_park, library, sev_day_avg_library, grocery, sev_day_avg_grocery, social_rec) %>% 
  group_by(prescribed_group) %>% 
  nest() %>% 
  mutate(
    model = map(data, function(d) plm(motivation ~ grocery, index = c("userId", "activityDay"), data = d, model = "within"))
  ) 

modelsummary(locations$model %>% set_names(str_c("Group: ", locations$prescribed_group)),
             gof_map = gm, coef_map = cm_comp,
             vcov = function(x) vcovHC(x, cluster = "time"),
             estimate = c("{estimate}{stars}"),
             statistic = c("({statistic})"),
             notes = list("Robust t-statistics in parentheses. + p < 0.1, * p < 0.05, ** p < 0.01, *** p < 0.001"),
             output = "kableExtra")
```

### Activity Impact on Motivation

These findings highlight the importance of considering not only how much individuals engage in activities, but also where those activities take place.
The varying effects of activity location on motivation across groups underscore the need for location-specific and group-sensitive strategies when evaluating the impact of travel behavior on mental well-being.
For the control group, the positive correlation between park visits and motivation suggests that outdoor environments may play a beneficial role in supporting overall well-being.
In contrast, the negative relationship between grocery store visits and motivation in the autism group points to the potential stress associated with routine, often overstimulating environments.

These group-specific patterns are important because they reveal how everyday activities can influence mental well-being differently depending on individual needs and diagnoses.
For example, promoting park visits may be particularly effective in enhancing well-being among the general population, while interventions aimed at reducing sensory stress in grocery store environments could support better outcomes for autistic individuals.
Notably, for individuals with social anxiety, the strong positive relationship between activity levels and motivation suggests that simply encouraging more engagement in daily activities, regardless of location, may be especially beneficial.
While specific settings may not significantly impact motivation in this group, the act of participating in activities itself appears to support mental well-being.
Together, these findings highlight the importance of tailoring mental health interventions to both the type of activity and the specific needs of different groups.
