# Results and Discussions

```{r setup2, file = "R/chapter_start.R", include = FALSE, cache = FALSE}
library(plm)
library(sandwich)
```

```{r model-setups, include=FALSE}
f <- function(x) format(round(x, 3), big.mark=",")
gm <- list(
  list("raw" = "nobs", "clean" = "Number of Observations", "fmt" = f),
  list("raw" = "logLik", "clean" = "Log. Liklihood", "fmt" = f),
  list("raw" = "aic", "clean" = "AIC", "fmt" = f),
  list("raw" = "r.squared", "clean" = "$R^2$", "fmt" = f))

cm <- c(
  "sev_day_avg" = "Seven-Day No. of Activities",
  "sexFemale" = "Female", 
  "age" = "Age", 
  "fsiq_2" = "IQ", 
  "prescribed_groupAutism" = "Autism", 
  "prescribed_groupSocial Anxiety" = "Social Anxiety", 
  "suicidal_ideation_q31_evenTRUE" = "Suicidal Ideation",
  "energy" = "Energy",
  "motivation" = "Motivation", 
  "area" = "Activity Area",
  "length" = "Distance", 
  "numTrips" = "No. of Activities")

cm_comp <- c(
  "numTrips" = "No. of Activities",
  "log(numTrips + 0.1)" = "log(No. of Activities)",
  "I(numTrips^2)" = "$No. of Activities^2$",
  "sev_day_avg" = "Seven-Day No. of Activities",
  "area" = "Activity Area",
  "length" = "Distance",
  "log(length)" = "log(Distance)",
  "I(length^2)" = "$Distance^2$")

cm_group <- c(
  "numTrips" = "No. of Activities",
  "sev_day_avg" = "Seven-Day No. of Activities",
  "park" = "Visits to Parks",
  "grocery" = "Visits to Grocery Store",
  "library" = "Visits to Libraries",
  "social_rec" = "Visits to Social Recreation Locations")
```

## Group Analysis

Analyzing travel behavior and mental well-being entails a comprehensive examination of the individual's group. It is become imperative to delve into various factors about how these groups travel and how their travel affects their mental well-being. Understanding the number and type of activities they engage in, motivational levels, and propensity towards suicidal ideation provides valuable insights for this analysis.

### Activity Engagement by Group

The analysis delved into the activity engagement patterns of the three distinct groups: control, autism, and social anxiety. Examining various facets of activity involvement, the findings offer valuable insights into the behavioral differences across these groups. We analyzed the total number of activities and the seven-day rolling average number of activities. The descriptive statistics in @tbl-groupActs describe these results for each group.

```{r table groupAct, echo=FALSE, message = FALSE}
#| label: tbl-groupActs
#| tbl-cap: "Activity Descriptive Statistics by Group"

tar_load(model_data)

acts <- model_data %>% 
  filter(!is.na(numTrips)) %>%
  select(prescribed_group, numTrips, sev_day_avg) %>% 
# , park, grocery, library, social_rec) %>% 
  group_by(prescribed_group) %>% 
  rename("No. of Activities" = numTrips,
         "Seven-Day No. of Activities" = sev_day_avg)

         # "Visits to Parks" = park,
         # "Visits to Grocery Store" = grocery,
         # "Visits to Libraries" = library,
         # "Visits to Social Recreation Locations" = social_rec)

datasummary_balance(~prescribed_group, data = acts, dinm=FALSE)
```

Starting with the total number of activities, individuals in the control group displayed a moderate level of engagement, as evidenced by a mean of 3.0 activities with a notable standard deviation of 2.7. This suggests a wide range of activity participation within the group. In contrast, participants in the autism group exhibited a lower mean of 2.1 activities, indicating comparatively fewer overall activities. However, the standard deviation of 2.3 suggests that while activity levels are generally lower than the control group, there is still considerable variability within the autism group. Similarly, individuals with social anxiety demonstrated a mean of 2.6 activities, falling between the control and autism groups. The standard deviation of 2.7 suggests variability similar to that observed in the control group. Moving to the seven-day rolling average of activities, trends mirrored those of the total number of activities, with slightly lower standard deviations, indicating less variability in the rolling average compared to the total count. We conducted an ANOVA test to see if the differences in the mean number of activities between the groups are statistically significant.

The ANOVA test was conducted to determine if significant differences exist in the mean number of activities among the three groups: control, autism, and social anxiety. The results from the ANOVA test are shown in @tbl-ANOVAActs and @fig-ANOVAActs shows the mean number of activities and 95% confidence intervals for the differences between the corresponding group means.

```{r table ANOVAAct, echo=FALSE, message = FALSE}
#| label: tbl-ANOVAActs
#| tbl-cap: "ANOVA Test: Number of Activity by Group"

library(xtable)

actsANOVA <- model_data %>% 
  filter(!is.na(numTrips)) %>%
  mutate(prescribed_group = as.factor(prescribed_group)) %>% 
  rename(Group = prescribed_group)

# Perform ANOVA to test for differences in mean number of activities among prescribed groups
ACTS_anova_result <- aov(numTrips ~ Group, data = actsANOVA)

summary(ACTS_anova_result)

# Summarize ANOVA results
# AA <- summary(ACTS_anova_result)
# xtable(AA)
```

```{r figure ANOVAActs, message=FALSE, echo=FALSE}
#| label: fig-ANOVAActs
#| fig-cap: "Mean number of activities by group."

ggplot(actsANOVA, aes(x = Group, y = numTrips)) +
  stat_summary(fun = mean, geom = "point", color = "black") +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2) +
  labs(x = "Group", y = "Number of Activities") +
  theme_bw()
```

The results unveiled a highly significant effect of the group on the number of activities, as evidenced by an F value of 32.97 and a p-value less than 0.001. This signifies that the variation in the number of activities across the groups is improbable to be attributed to random chance, confirming the presence of substantial differences among the groups.

After determining that there are significant differences in the mean number of activities among three groups, a Tukey's Honest Significant Difference (HSD) test was conducted to pinpoint which specific groups differ from each other. The results are shown in @tbl-TukeyActs.

```{r table TukeyHSD-acts, message=FALSE, echo=FALSE}
#| label: tbl-TukeyActs
#| tbl-cap: "TukeyHSD Test: Difference in Mean Motivation Across Groups"

# Perform Tukey's HSD test for pairwise comparisons
TukeyHSD(ACTS_anova_result)
```

Tukey's HSD test indicated that there were significant differences in mean number of activities between all pairs of prescribed groups. Specifically, individuals in the autism group engaged in significantly fewer activities on average compared to those in the control group, with a mean difference of -0.896 and a p-value less than 0.001, and the social anxiety group, with a mean difference of -0.375 and a p-value less than 0.001. Similarly, participants in the social anxiety group participated in significantly fewer activities on average than those in the control group with a mean difference of -0.521 and a p-value less than 0.001.

These findings suggest that there are significant differences in the mean number of activities among individuals belonging to different groups. Specifically, individuals with autism tend to engage in fewer activities compared to both the control group and the social anxiety group. Similarly, individuals in the social anxiety group participate in fewer activities on average compared to those in the control group. These results highlight the importance of considering group differences when examining activity engagement patterns and potentially mental health outcomes.

### Motivation by Group

Upon closer examination of the distinct groups, notable differences in well-being also emerge. Drawing from existing literature, it was anticipated that individuals in both the autism and social anxiety groups would exhibit lower levels of well-being compared to those in the control group. Given our use of motivation as an indicator of well-being, we analyzed the reported levels of motivation across each group. @tbl-groupMotiv presents the mean and standard deviation of motivation for each group.

```{r table groupMotiv, message=FALSE, echo=FALSE}
#| label: tbl-groupMotiv
#| tbl-cap: "Motivation Levels by Group"

tar_load(model_data)

motiv <- model_data %>% 
  filter(!is.na(numTrips)) %>%
  select(prescribed_group, motivation) %>% 
  group_by(prescribed_group) %>% 
  mutate("Level of Motivation" = motivation)

motiv_tbl <- motiv %>% 
  select(prescribed_group, "Level of Motivation")

datasummary_balance(~prescribed_group, data = motiv_tbl, dinm=FALSE)
```

The findings revealed differences in motivation levels across these groups. The control group had the highest mean motivation score of 47.7, with a standard deviation of 14.2. This suggests that individuals in the control group generally have high motivation levels with relatively little variability, indicating a consistent trend of high motivation within this group. In contrast, the autism group showed the lowest mean motivation score of 34.0 and the highest standard deviation of 19.4. The lower mean score implies that individuals with autism tend to have lower motivation levels compared to those in the control group. Additionally, the high standard deviation indicates significant variability in motivation among participants in this group, suggesting that motivation levels are more dispersed. The social anxiety group had a mean motivation score of 40.3 and a standard deviation of 18.5. This group's motivation levels are higher than those of the autism group but lower than the control group's. The standard deviation, while high, is slightly lower than that of the autism group, indicating considerable but somewhat less variability in motivation levels compared to the autism group. To determine if the differences in mean motivation levels are significantly different, we conducted an ANOVA test.

The ANOVA test was conducted to determine if there are significant differences in the mean levels of motivation among three groups: control, autism, and social anxiety. The results revealed a highly significant effect of the prescribed group on motivation levels, with an F value of 362 and a p-value less than 0.001. This indicates that the variation in motivation levels across the prescribed groups is unlikely to be due to random chance, and there are indeed significant differences among the groups. The results are summarized in @tbl-ANOVAMotiv and @fig-ANOVAMotiv shows the mean level of motivation and 95% confidence intervals for the differences between the corresponding group means.

```{r table ANOVA-motivation, message=FALSE, echo=FALSE}
#| label: tbl-ANOVAMotiv
#| tbl-cap: "ANOVA Test: Difference in Mean Motivation Across Groups"

motivANOVA <- model_data %>% 
  filter(!is.na(motivation))%>%
  mutate(prescribed_group = as.factor(prescribed_group)) %>% 
  rename(Group = prescribed_group)

# Perform ANOVA
MOTIV_anova_result <- aov(motivation ~ Group, data = motivANOVA)

summary(MOTIV_anova_result)

# Print the ANOVA table
# AM <- summary(MOTIV_anova_result)
# xtable(AM)
```

```{r figure ANOVAMotiv, message=FALSE, echo=FALSE}
#| label: fig-ANOVAMotiv
#| fig-cap: "Mean levels of motivation by group."

ggplot(motivANOVA, aes(x = Group, y = motivation)) +
  stat_summary(fun = mean, geom = "point", color = "black") +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.2) +
  labs(x = "Group", y = "Motivation") +
  theme_bw()
```

After determining that there significant differences in the mean levels of motivation among three groups, a Tukey's Honest Significant Difference (HSD) test was conducted to pinpoint which specific groups differ from each other. The results are shown in @tbl-TukeyMotiv.

```{r table TukeyHSD-motivation, message=FALSE, echo=FALSE}
#| label: tbl-TukeyMotiv
#| tbl-cap: "TukeyHSD Test: Difference in Mean Motivation Across Groups"

TukeyHSD(MOTIV_anova_result)
```

The autism group had a significantly lower mean motivation level compared to the control group, with a difference of -9.99 points. This difference was highly significant with a p-value of less than 0.001. Similarly, the social anxiety group also exhibited a significantly lower mean motivation level compared to the control group, with a difference of -7.74 points and a p-value less than 0.001. When comparing the social anxiety group to the autism group, the social anxiety group showed a significantly higher mean motivation level by 2.25 points, with this difference also being statistically significant, p-value less than 0.001.

Overall, the results of the ANOVA test indicate that there are significant differences in mean motivation levels among the three prescribed groups. Specifically, both the autism and social anxiety groups have significantly lower motivation levels compared to the control group. Moreover, within the groups that experienced lower motivation, the social anxiety group had significantly higher motivation levels than the autism group.

These findings suggest that the group has a significant influence on motivation levels. The control group exhibited the highest mean motivation levels, followed by the social anxiety group, and then the autism group, which had the lowest mean motivation levels. This pattern underscores the impact that the type of prescribed group may have on an individual's motivation, highlighting the need for further exploration into the association between travel behavior and well-being.

### Suicidal Ideation by Group

The morning and evening surveys included different questions pertaining to suicidal ideation. Specifically, the evening survey delved into various aspects of suicidality, asking participants with the following questions:

-   Have you thought about killing yourself in the past 12 hours or since you last took a survey?

-   How intense was your desire to kill yourself since the last survey you completed (or over the past 12 hours, if you didn't complete the last survey)?

-   How strong was your intention to kill yourself by suicide since the last survey you completed (or over the past 12 hours, if you didn't complete the last survey)?

-   How strong was your ability to resist the urge to kill yourself since the last survey you completed (or over the past 12 hours, if you didn't complete the last survey)?

The dataset presents an insightful glimpse into the prevalence of suicidal ideation within distinct demographic groups, shedding light on potential differences in mental health concerns among them. Examining responses to the question "Have you thought about killing yourself in the past 12 hours or since you last took a survey?" across the three groups: control, autism, and social anxiety, unveils intriguing patterns. The responses to this question are summarized by group in @tbl-groupSuicide.

```{r table groupSuicide, message=FALSE, echo=FALSE}
#| label: tbl-groupSuicide
#| tbl-cap: "Suicidal Ideation by Group"

tar_load(model_data)

suic <- model_data %>% 
  filter(!is.na(numTrips)) %>%
  select(prescribed_group, suicidal_ideation_q31_even) %>%
  mutate(suicidal_ideation_q31_even = case_when(
    suicidal_ideation_q31_even == TRUE ~ "Yes",
    suicidal_ideation_q31_even == FALSE ~ "No",
    is.na(suicidal_ideation_q31_even) ~ "No Response",
    TRUE ~ as.character(suicidal_ideation_q31_even)
  )) %>%
  mutate(suicidal_ideation_q31_even = factor(suicidal_ideation_q31_even, levels = c("Yes", "No", "No Response"))) %>%
  group_by(prescribed_group) %>% 
  rename("Suicidal Ideation" = suicidal_ideation_q31_even)

datasummary_balance(~prescribed_group, data = suic, coef_map = cm_group, dinm=FALSE)
```

In the control group, on 66.4% of days participants responded negatively, indicating they did not think about killing themselves. However, on 2.1% of days, group members acknowledged experiencing suicidal ideation during the specified timeframe. Contrastingly, among respondents with autism, on 51.6% of the days, participants denied having suicidal thoughts, while respondents admitted to suicidal ideation with a slightly higher percentage on 4.9% of days. In the social anxiety group, the dynamics were notably different. Here, participants reported no suicidal ideation on a significantly lower percentage, 38.9%, of days, while they reported suicidal ideation on a strikingly higher proportion of days, at 15.6% of days. Across all three groups, there was a significant portion of respondents that chose not to provide a response, perhaps indicating the sensitivity of the question or reluctance to disclose personal struggles.

To dive deeper into the suicidal ideation of individuals in the different groups, we constructed a contingency table. The contingency table organized the data into rows and columns, with each row representing a group (control, autism, social anxiety) and each column representing a response category ("Yes," "No," "No Response") regarding suicidal ideation. Within the table, the counts of individual responses in each group were recorded. Subsequently, a chi-square test of independence was conducted on these counts to assess whether there is a significant association between the prescribed group and the response to suicidal ideation. The results of the chi-square test are shown in @tbl-groupSuicideChi.

```{r table groupSuicideChi, message=FALSE, echo=FALSE}
#| label: tbl-groupSuicideChi
#| tbl-cap: "Pearson Chi Test: Suicidal Ideation by Group"

contingency_table <- table(suic$prescribed_group, suic$`Suicidal Ideation`)

chisq_result <- chisq.test(contingency_table)

print(chisq_result)
```

The test calculated a chi-square statistic of 388.06 with 4 degrees of freedom. The resulting p-value was found to be less than 0.001, signifying substantial evidence against the null hypothesis of no association. Thus, the analysis concludes that there is a statistically significant association between the group and the response to suicidal ideation. This suggests that the group status influences the likelihood of a particular response to suicidal ideation, indicating potential differences in how individuals from different groups perceive and experience suicidal ideation. These findings offer valuable insights into the interplay between group and response to suicidal ideation.

## Model Comparison and Evaluation

As discussed previously, we ran three different models to analyze the affect of the seven-day rolling average number of activities on motivation levels. We ran the OLS, FE, and RE models with robust standard errors and t-statistics due to the autocorrelation and heteroskedasticity. The results of these three models are shown in @tbl-olsfere.

```{r table ols fe re, message=FALSE, echo=FALSE}
#| label: tbl-olsfere
#| tbl-cap: "OLS, FE, and RE Models"
#| warnings: FALSE
 
library(plm) 

tar_load(ols)
tar_load(fe)
tar_load(re)

models_olsfere <- list(
    "Ordinary Least Squares Model" = ols,
    "Fixed Effects Model" = fe,
    "Random Effects Model" = re
   )

modelsummary(models_olsfere, gof_map = gm, coef_map = cm,
              vcov = function(x) vcovHC(x, cluster = "time"),
              estimate = c("{estimate} ({statistic}){stars}"),
              statistic = NULL,
              notes = list("Robust t-statistics in parentheses")
   )
```

We used the Hausman test to determine whether the RE model is more appropriate than the FE model. It tests whether the RE estimates are consistent and efficient compared to the FE estimates. The null hypothesis is that the RE estimates are consistent and efficient, while the alternative hypothesis is that the FE estimates are more efficient. We found, using the Hausman test, that the p-value is 0.0013 which is less than 0.05, which means that the results are significant and we reject the null hypothesis. This implies that the coefficients from the FE model and RE model are sufficiently different from each other, which means that the RE model is inconsistent and the FE model should be used in the analysis.

## Effect of Demographic Factors on Motivation

Based on the results of the Hausman test, we found that the FE model should be used in the analysis of the activity pattern and mental health data. One potential downside of the The FE model is that it cannot implement time constant variables. To overcome this limitations, we performed a linear regression analysis to examine how demographic variables (e.g., sex, age, IQ, and group) are associated with the intercepts from the FE model. This allows for understanding of how the baseline levels of motivation differ across different demographic groups. The sex, age, IQ, and group were the independent variables, and the FE intercept values for each userID served as the dependent variable. This model is shown in @tbl-fedemolm.

```{r table fe demo lm, message=FALSE, echo=FALSE}
#| label: tbl-fedemolm
#| tbl-cap: "Fixed Effects and Demographics Regression"

tar_load(fe_model)

modelsummary(fe_model, gof_map = gm, coef_map = cm,
             estimate = c("{estimate} ({statistic}){stars}"),
             statistic = NULL,
             notes = list("t-statistics in parentheses"))
```

The linear model examined the relationship between motivation and various predictor variables, including sex, age, IQ, and group. The analysis revealed several significant findings. First, being female was associated with a decrease in motivation by 6.351 points, on a scale from 0-100, compared to males, which is a statistically significant effect with a p-values less than 0.05. However, age and IQ did not show significant associations with motivation. Conversely, individuals with autism exhibited a substantial decrease in motivation by 10.247 points compared to the control group, and those with social anxiety experienced a similar decrease of 8.544 points, both statistically significant at p less than 0.01. These results suggest that sex and group conditions play significant roles in shaping motivational levels, while age and IQ appear to have limited influence in this context. When looking at the fit of this model to the data, the model explains approximately 26.4% of the variance in motivation, indicating that other unexplored factors may contribute to motivational outcomes. These findings underscore the importance of considering individual differences, particularly sex and group status, when examining levels of motivation.

## Model Comparison for Impact on Motivation

To compare the effects of different aspects of travel behavior on motivation, we conducted a comprehensive analysis considering various factors. Specifically, we examined the impact of the total number of activities, the seven-day rolling average number of activities, the convex hull area, and the total distance traveled on a day.

Analyzing the total number of activities provides insights into the overall level of engagement and mobility of individuals. A higher number of activities may indicate greater involvement in social, recreational, or productive tasks, which could positively result in an increase in motivation. Similarly, a lower number of activities might suggest reduced participation and potential disengagement, possibly leading to decreased motivation.

The seven-day rolling average number of activities allows us to assess the consistency and stability of individuals' activity patterns over time. Fluctuations in this metric may reflect changes in routine, lifestyle adjustments, or shifts in mental well-being, all of which can influence motivation. A higher rolling average may indicate sustained engagement and stability, potentially contributing to higher motivation levels.

The convex hull area represents the spatial extent covered by an individual's activities, providing insights into their travel behavior's spatial distribution and range. A larger convex hull area may suggest greater exploration, mobility, and exposure to diverse environments, which could positively impact motivation through experiences and opportunities. Conversely, a larger convex hull area may suggest a lot of commuting which may lead to a decrease in motivation.

Lastly, the total distance traveled on a day captures the extent of physical movement and travel undertaken by individuals. Longer distances traveled may indicate higher levels of physical activity, exploration, or commuting, all of which can impact energy levels and motivation. Understanding how travel distance relates to motivation can provide valuable insights into the interplay between physical activity and mental well-being.

By examining these different aspects of travel behavior in relation to motivation, we aim to discern which factors have the most significant influence and in what ways. This holistic approach will enable us to gain a comprehensive understanding of how travel behavior interacts with motivation, offering valuable insights for promoting mental well-being and enhancing motivation levels among individuals.

The analysis examined how motivation is influenced by different independent variables across four distinct models: the Number of Activities Model, the Seven-Day Rolling Average Number of Activities Model, the Activity Area Model, and the Distance Traveled Model. Motivation served as the dependent variable in each of these models. The results from these models are shown in @tbl-motivimpact.

```{r table motivation impact, message=FALSE, echo=FALSE}
#| label: tbl-motivimpact
#| tbl-cap: "Comparison of Impacts on Motivation Levels"

tar_load(fe_sevdayavg_mod)
tar_load(fe_numTrips_mod)
tar_load(fe_area_mod)
tar_load(fe_length_mod)

models_fe_paras <- list(
  "I" = fe_numTrips_mod,
  "II" = fe_sevdayavg_mod,
  "III" = fe_area_mod,
  "IV" = fe_length_mod
  )

modelsummary(models_fe_paras, gof_map = gm, coef_map = cm_comp, 
             fmt = fmt_sprintf("%.3e"), 
             vcov = function(x) vcovHC(x, cluster = "time"),
             estimate = c("{estimate} ({statistic}){stars}"),
             statistic = NULL,
             notes = list("Robust t-statistics in parentheses")
             )
```

The findings from these models suggest that there is a positive relationship between the number of activities and motivation and the relationship is statistically significant. For each additional activity a person engages in, their motivation increases 0.356 points. While motivation is on a scale from 1-100, it may not be reasonable to engage is one more activity in a day to increase motivation by 0.356 points, especially considering that the average number of activities for all of the userID-activity days was 2.65 activities.

The results also suggest that there is a negative relationship between the distance traveled and the motivation; this relationship is also statistically significant. In the case of distance traveled, for each additional kilometer traveled, the motivation of the individual decreased by 0.0003263 points. Similar to the case with the number of activities, while the distance traveled is statistically significant, the decrease in motivation seems so minor that it would not be reasonable to use distance traveled to inform changes in motivation.

The seven-day rolling average of activities also shows a potential cumulative effect on motivation, though the evidence is weaker. In contrast, the size of the activity area does not appear to influence motivation, indicating that the physical space covered during the day is less important than the number of activities themselves. Additionally, traveling longer distances during the day might slightly decrease motivation.

Overall, while the number of activities, seven-day rolling average number of activities, and distance traveled are important factor for motivation, the low R-squared values across all models suggest that many other factors not included in this analysis likely play a significant role in determining motivation levels.

### Number of Activities Models

To further analyze the impact of the number of activities, we applied transformations to the independent variable. We looked at the logarithm of the number of activities and the number of activities squared. @tbl-motivnumActs compares the original model with the number of trips with these two transformed models.

```{r table motivation numActs, message=FALSE, echo=FALSE}
#| label: tbl-motivnumActs
#| tbl-cap: "Comparison of the Number of Activities Models"

tar_load(fe_numTrips_mod)
tar_load(fe_numTrips_log)
tar_load(fe_numTrips_squared)

models_numActs <- list(
  "I" = fe_numTrips_mod,
  "II" = fe_numTrips_log,
  "III" = fe_numTrips_squared
  )

modelsummary(models_numActs, gof_map = gm, coef_map = cm_comp,
             vcov = function(x) vcovHC(x, cluster = "time"),
             estimate = c("{estimate} ({statistic}){stars}"),
             statistic = NULL,
             notes = list("Robust t-statistics in parentheses")
             )
```

### Activity Area Covered

To further analyze the impact of the distance traveled, we applied transformations to the independent variable. We looked at the logarithm of the distance traveled and the distance traveled squared. @tbl-motivDistTrav compares the original model with the distance traveled with these two transformed models.

```{r table motivation numActs, message=FALSE, echo=FALSE}
#| label: tbl-motivDistTrav
#| tbl-cap: "Comparison of the Distance Traveled Models"

tar_load(fe_length_mod)
tar_load(fe_length_log)
tar_load(fe_length_squared)

models_distTrav <- list(
  "I" = fe_length_mod,
  "II" = fe_length_log,
  "III" = fe_length_squared
  )

modelsummary(models_distTrav, gof_map = gm, coef_map = cm_comp,
             fmt = fmt_sprintf("%.3e"),
             vcov = function(x) vcovHC(x, cluster = "time"),
             estimate = c("{estimate} ({statistic}){stars}"),
             statistic = NULL,
             notes = list("Robust t-statistics in parentheses")
             )
```

