# Discussion and Applications {#sec-contribs}

```{r setup3, file = "R/chapter_start.R", include = FALSE, cache = FALSE}

```

```{r model-setups, include=FALSE}
f <- function(x) format(round(x, 3), big.mark=",")
gm <- list(
  list("raw" = "Num.Obs.", "clean" = "Number of Observations", "fmt" = f),
  list("raw" = "Log.Lik.", "clean" = "Log Liklihood", "fmt" = f),
  list("raw" = "rho20", "clean" = "$\\rho^2_0$", "fmt" = f))
  
cm_group <- c(
  "numTrips" = "Number of Activities",
  "sev_day_avg" = "7-Day Rolling Avg. Number of Activities",
  "park" = "Visits to Parks",
  "grocery" = "Visits to Grocery Store",
  "library" = "Visits to Libraries",
  "social_rec" = "Visits to Social Recreation Locations")
```

## Group Analysis

Analyzing travel behavior and mental well-being entails a comprehensive examination of the individual's group. Given the forthcoming exploration into how these groups travel and the impact of mental well-being on their travel patterns, it becomes imperative to delve into various factors. Understanding the number and type of activities they engage in, motivational levels, and propensity towards suicidal ideation provides valuable insights for this analysis.

### Activity Engagement by Group

The analysis delved into the activity engagement patterns of the three distinct groups: control, autism, and social anxiety. Examining various facets of activity involvement, the findings offer valuable insights into the behavioral differences across these groups. We analyzed the total number of activities, the seven-day rolling average number of activities, and activities at parks, grocery stores, libraries, and social recreation locations for each group. The descriptive statistics in @tbl-groupActs describe these results for each group.

```{r table groupAct, echo=FALSE, message = FALSE}
#| label: tbl-groupActs
#| tbl-cap: "Activity Descriptive Statistics by Group"

tar_load(model_data)

acts <- model_data %>% 
  filter(!is.na(numTrips)) %>%
  select(prescribed_group, numTrips, sev_day_avg, park, grocery, library, social_rec) %>% 
  group_by(prescribed_group) %>% 
  rename("Number of Activities" = numTrips,
         "7-day Rolling Avg. Number of Activities" = sev_day_avg,
         "Visits to Parks" = park,
         "Visits to Grocery Store" = grocery,
         "Visits to Libraries" = library,
         "Visits to Social Recreation Locations" = social_rec)

datasummary_balance(~prescribed_group, data = acts, dinm=FALSE)
```

Starting with the total number of activities, individuals in the control group displayed a moderate level of engagement, as evidenced by a mean of 3.0 activities with a notable standard deviation of 2.7. This suggests a wide range of activity participation within the group. In contrast, participants in the autism group exhibited a lower mean of 2.1 activities, indicating comparatively fewer overall activities. However, the standard deviation of 2.3 suggests that while activity levels are generally lower than the control group, there is still considerable variability within the autism group. Similarly, individuals with social anxiety demonstrated a mean of 2.6 activities, falling between the control and autism groups. The standard deviation of 2.7 suggests variability similar to that observed in the control group. Moving to the seven-day rolling average of activities, trends mirrored those of the total number of activities, with slightly lower standard deviations, indicating less variability in the rolling average compared to the total count.

In terms of activities at specific locations, engagement at parks, grocery stores, libraries, and social recreation locations was generally infrequent across all groups. For instance, activities at parks yielded minimal mean values and low standard deviations, suggesting sporadic engagement. Similarly, activities at grocery stores and libraries were rare, with mean values close to zero and minimal variability. Notably, individuals with social anxiety exhibited slightly higher mean values and greater variability in activities at grocery stores compared to other groups. Activities at social recreation locations, while still infrequent, showed slightly higher mean values across all groups, with some variability in engagement, particularly among individuals with control and social anxiety.

In summary, while individuals in the control group demonstrated a moderate level of activity engagement with considerable variability, those with autism and social anxiety exhibited lower overall activity engagement levels. Engagement at specific locations was minimal across all groups when looking at aggregated activity engagement for each group.

### Motivation by Group

Upon closer examination of the distinct groups, notable differences in well-being emerge. Drawing from existing literature, it is anticipated that individuals in both the autism and social anxiety groups would exhibit lower levels of well-being compared to those in the control group. Given our utilization of motivation as an indicator of well-being, we will assess the reported levels of motivation across each group. @tbl-groupMotiv presents the mean and standard deviation of motivation for each group, while @fig-groupMotiv illustrates the distribution of motivation across the groups.

```{r table groupMotiv, message=FALSE, echo=FALSE}
#| label: tbl-groupMotiv
#| tbl-cap: "Motivation Levels by Group"

tar_load(model_data)

motiv <- model_data %>% 
  filter(!is.na(numTrips)) %>%
  select(prescribed_group, motivation) %>% 
  group_by(prescribed_group) %>% 
  mutate("Level of Motivation" = motivation)

motiv_tbl <- motiv %>% 
  select(prescribed_group, "Level of Motivation")

datasummary_balance(~prescribed_group, data = motiv_tbl, dinm=FALSE)
```

```{r figure groupMotiv, message=FALSE, echo=FALSE}
#| label: fig-groupMotiv
#| fig-cap: "Distribution of motivation levels by group."

motiv_fig <- motiv %>% 
  select(prescribed_group, motivation)

ggplot(motiv_fig, aes(x = prescribed_group, y = motivation, fill = prescribed_group)) +
  geom_boxplot(fill = "darkgray", color = "black") +
  theme_bw() +
  labs(x = "Group", y = "Level of Motivation")
```

The findings revealed notable differences in motivation levels across these groups. The control group, comprising of 1,706 responses, had the highest mean motivation score of 47.7, with a standard deviation of 14.2. This suggests that individuals in the control group generally have high motivation levels with relatively little variability, indicating a consistent trend of high motivation within this group. In contrast, the autism group, with 804 responses, showed the lowest mean motivation score of 34.0 and the highest standard deviation of 19.4. The lower mean score implies that individuals with autism tend to have lower motivation levels compared to those in the control group. Additionally, the high standard deviation indicates significant variability in motivation among participants in this group, suggesting that motivation levels are more dispersed. The social anxiety group, consisting of 1,893 responses, had a mean motivation score of 40.3 and a standard deviation of 18.5. This group's motivation levels are higher than those of the autism group but lower than the control group's. The standard deviation, while high, is slightly lower than that of the autism group, indicating considerable but somewhat less variability in motivation levels compared to the autism group.

Overall, these results highlight distinct differences in motivation levels between the groups. The control group exhibits the highest and most consistent motivation levels. In contrast, the autism group shows the lowest motivation levels with significant variability, and the social anxiety group falls in between, with moderate motivation levels and substantial variability. These findings suggest that both the condition, autism or social anxiety, and the variability within these groups are important factors in understanding motivation and thus, well-being.

### Suicidal Ideation by Group

The morning and evening surveys included different questions pertaining to suicidal ideation. Specifically, the evening survey delved into various aspects of suicidality, asking participants with the following questions:

-   Have you thought about killing yourself in the past 12 hours or since you last took a survey?

-   How intense was your desire to kill yourself since the last survey you completed (or over the past 12 hours, if you didn't complete the last survey)?

-   How strong was your intention to kill yourself by suicide since the last survey you completed (or over the past 12 hours, if you didn't complete the last survey)?

-   How strong was your ability to resist the urge to kill yourself since the last survey you completed (or over the past 12 hours, if you didn't complete the last survey)?

The dataset presents an insightful glimpse into the prevalence of suicidal ideation within distinct demographic groups, shedding light on potential differences in mental health concerns among them. Examining responses to the question "Have you thought about killing yourself in the past 12 hours or since you last took a survey?" across the three groups—control, autism, and social anxiety, unveils intriguing patterns. The responses to this question are summarized by group in @tbl-groupSuicide.

```{r table groupSuicide, message=FALSE, echo=FALSE}
#| label: tbl-groupSuicide
#| tbl-cap: "Suicidal Ideation by Group"

tar_load(model_data)

suic <- model_data %>% 
  filter(!is.na(numTrips)) %>%
  select(prescribed_group, suicidal_ideation_q31_even) %>%
  mutate(suicidal_ideation_q31_even = case_when(
    suicidal_ideation_q31_even == TRUE ~ "Yes",
    suicidal_ideation_q31_even == FALSE ~ "No",
    is.na(suicidal_ideation_q31_even) ~ "No Response",
    TRUE ~ as.character(suicidal_ideation_q31_even)  # for any other cases
  )) %>%
  mutate(suicidal_ideation_q31_even = factor(suicidal_ideation_q31_even, levels = c("Yes", "No", "No Response"))) %>%
  group_by(prescribed_group) %>% 
  rename("Suicidal Ideation" = suicidal_ideation_q31_even)

datasummary_balance(~prescribed_group, data = suic, coef_map = cm_group, dinm=FALSE)
```

In the control group, on 66.4% of days participants responded negatively, indicating they did not think about killing themselves. However, on 2.1% of days, group members acknowledged experiencing suicidal ideation during the specified timeframe. Contrastingly, among respondents with autism, on 51.6% of the days, participants denied having suicidal thoughts, while respondents admitted to suicidal ideation with a slightly higher percentage on 4.9% of days. In the social anxiety group, the dynamics were notably different. Here, participants reported no suicidal ideation on a significantly lower percentage, 38.9%, of days, while they reported suicidal ideation on a strikingly higher proportion of days, at 15.6% of days.

Across all three groups, there was a significant portion of respondents that chose not to provide a response, perhaps indicating the sensitivity of the question or reluctance to disclose personal struggles.

These findings suggest nuanced variations in the prevalence of suicidal ideation across the groups, underscoring the importance of tailored interventions and support mechanisms to address the diverse mental health needs within these populations.

## Relationship between suicidal ideation and motivation

Considering the data presented on the three groups—control, social anxiety, and autism—several potential connections between their activity engagements, motivation levels, and suicidal ideation can be explored.

Activity engagements may serve as a reflection of one's mental state and well-being. The differences observed in the number of activities and their nature across the groups could indicate variations in their daily routines, coping mechanisms, and social interactions. For instance, individuals in the control group might exhibit higher activity engagement, possibly indicative of better mental health and social integration compared to those in the social anxiety and autism groups.

Motivation levels could play a significant role in influencing activity engagements and, consequently, mental well-being. Lower motivation levels, as observed in the autism group and social anxiety group compared to the control group, might contribute to decreased participation in activities and decreased overall satisfaction with life. This, in turn, could exacerbate feelings of isolation, loneliness, and low self-esteem, which are known risk factors for suicidal ideation.

The prevalence of suicidal ideation across the groups underscores the complex interplay between mental health and activity engagement. Individuals experiencing higher levels of suicidal ideation may exhibit decreased motivation to engage in activities and may withdraw from social interactions, leading to further isolation and worsening mental health outcomes. Conversely, active engagement in meaningful activities and supportive social networks could serve as protective factors against suicidal ideation. In addition, the literature suggests that suicidal behavior is negatively associated with overall well-being [@fonseca-pedreroRiskProtectiveFactors2022; @fumeroAdolescentsBipolarExperiences2021]. Since we connected motivation to well-being, a similar association is drawn between suicidal tendency and motivation.

Overall, the data suggest potential links between activity engagements, motivation levels, and suicidal ideation among the three groups.

@fig-groupAnaly shows the relationship between motivation and suicidal intensity by group.

```{r figure groupAnaly, message=FALSE, echo=FALSE}
#| label: fig-groupAnaly
#| fig-cap: "Motivation vs Suicidal Intensity"

tar_load(model_data)

analysis <- model_data %>%
  select(userId, prescribed_group, activityDay, motivation, numTrips, sev_day_avg, suicidal_intensity_q32_even) %>% 
  filter(!is.na(suicidal_intensity_q32_even) & suicidal_intensity_q32_even != "CONDITION_SKIPPED") %>% 
  mutate(suic_intensity = as.numeric(suicidal_intensity_q32_even)) %>% 
  filter(!is.na(suic_intensity))

ggplot(analysis, aes(y = motivation, x = suic_intensity, color = prescribed_group)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  facet_wrap(~prescribed_group) +
  labs(y = "Level of Motivation", x = "Suicidal Intensity") +
  theme_bw()
```

@fig-groupAnaly2 shows the relationship between motivation and the number of activities by group.

```{r figure groupAnaly2, message=FALSE, echo=FALSE}
#| label: fig-groupAnaly2
#| fig-cap: "Motivation vs Number of Activities"

ggplot(analysis, aes(y = motivation, x = numTrips, color = prescribed_group)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  facet_wrap(~prescribed_group) +
  labs(y = "Level of Motivation", x = "Number of Activities") +
  theme_bw()
```

@fig-groupAnaly3 shows the relationship between suicidal intesity and the number of activities by group.

```{r figure groupAnaly3, message=FALSE, echo=FALSE}
#| label: fig-groupAnaly3
#| fig-cap: "Number of Activities vs Suicidal Intensity"

ggplot(analysis, aes(y = suic_intensity, x = numTrips, color = prescribed_group)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  facet_wrap(~prescribed_group) +
  labs(y = "Suicide Intensity", x = "Number of Activities") +
  theme_bw()
```